\documentclass[main.tex]{subfiles}

\begin{document}

\section{Experimental design}
\subsection{Setting}
We have two sets of N2G models of individual neurons or SAE latents, 
and a set of text samples for each N2G model.
We also decide on a firing threshold that determines 
whether a neuron or feature is considered active for a given token.
\subsection{Goal}
There are two main goals of this experiment:
\begin{enumerate}
    \item To measure the performance of individual N2G models 
    quantatively with uncertainty estimates.
    This is useful in order to know 
    whether the N2G model is reliable for each specific neuron/feature.
    \item To compare the overall performance of the two sets of N2G models.
    This allows us to compare different ways of creating N2G models 
    and to compare the performance of N2G models of individual neurons 
    to those of SAE latents.
\end{enumerate}
\subsection{Procedure}
The procedure for each goal is as follows:
\begin{enumerate}
    \item For each N2G model, 
    we obtain predicted activations for all tokens in the text samples 
    and obtain the ground truth by running the original language model 
    on the same text samples.
    We now have a set of binary predictions and ground truths 
    for a number of tokens.
    From this we can calculate the precision, recall, and F1 score.
    \item I am more uncertain about how to achieve this goal.
    Currently I just take the average statistics over all N2G models 
    in each set.
    We could also see each set as a single model for predicting 
    the activations of all neurons/latents 
    and calculate a single set of statistics for each set.
\end{enumerate}
We currently use \texttt{sklearn.metrics} to calculate the statistics 
and this does not give us uncertainty estimates.
\subsection{Discussion}
\begin{itemize}
    \item Currently, we use 16 text samples for training each N2G model 
    and 16 for testing.
    This is not many for either purpose.
    The issue is that the samples we use are max activating samples 
    and acquiring these is expensive, especially in disk usage.
    We could probably scale up to 64 samples or so, 
    at least for small models.
    Therefore it would be beneficial to reduce our reliance on these.
    An idea for doing this is to use some other samples for either 
    training or testing.
    How do we test whether this works?
    \item As described above, we use a fixed firing threshold 
    to turn continuous activations into binary predictions.
    Since both the predictions and the ground truth 
    are originally continuous, 
    we could instead compare these values directly.
    This would remove an arbitrary parameter (the firing threshold) 
    as long with allowing more fine grained comparisons.
    The issue is that it is not clear that there is any useful information 
    in the exact prediction values from the N2G models.
\end{itemize}


\nocite{foote_neuron_2023}
\subbib{main}
\end{document}