%!LW recipe=pdflatex
\documentclass[main.tex]{subfiles}

\begin{document}
GitHub: \url{https://github.com/albertsgarde/thesis}
Newest version of report: \url{https://albertsgarde.github.io/thesis/main/report.pdf}

I see running N2G on some SAE as a MVP of my thesis.
After that, the only core thing is to train my own SAEs and run N2G on them.
Both these should be doable within the next month (assuming non-extreme amounts of inteference).
Once these are done it'll be time to start looking into all the stuff we've talked about with measuring the interpretability of the graphs and stuff while I continue work on the report.

\section*{Progress since last meeting}
Overall not nearly as much as I could have wanted because of the revy and some unexpected personal stuff.
At least the revy is now over so I can focus on the thesis again.
\begin{itemize}
    \item Reworked the N2G code to be ready for SAEs.
    \item Did some reading in the N2G paper.
    \item Started work on getting SAEs from \verb|jacobcd52/gpt2-small-sparse-autoencoders| on HuggingFace to work.
    Seems pretty doable.
\end{itemize}
\section*{Plan for the next weeks}
\begin{itemize}
    \item Run N2G on at least one feature of at least one SAE.
    \item Start work on training own SAEs.
    \item Write theory sections on N2G and MAS.
    Theory on SAEs should be easier once I've trained my own.
    \item Probably do some reading of other papers?
\end{itemize}
\section*{Topics for meeting}
\begin{itemize}
    \item The issue with too many ECTS should be resolved now, so I can officially start.
\end{itemize}

\end{document}
