\documentclass[aspectratio=169]{beamer}
%\setbeameroption{show notes}

\usepackage{beamer_pre}


\title{Thesis meeting 2024/03/27}
\author{Albert R. S. Garde}
\date{\today}
	

\begin{document}

\frame{
	\maketitle
}

\begin{frame}[fragile=singleslide]
	\frametitle{Progress since last meeting}
    \begin{itemize}
        \item Found critical issues with the results from last week when looking at MAS samples.
        \item Created Rust implementation of the feature model in N2G.
        \item Tried to get my new \verb|prune| implementation to work.
        \item Looked at some other interpretability measures.
    \end{itemize}
\end{frame}
\begin{frame}[fragile=singleslide]
    \frametitle{Plan for the next weeks}
    \begin{itemize}
        \item Fix the critical issue.
        \item Redo the statistical analysis 
        (maybe with the bootstrap CI test instead of the Chebyshev inequality) 
        and again take a look at the individual feature models and MAS samples.
        \item Write draft of methodology section.
        \item Integrate the Rust implementation of the feature model in N2G.
    \end{itemize}
\end{frame}
\begin{frame}[fragile=singleslide]
    \frametitle{Topics for meeting}
    \begin{itemize}
        \item Show examples of issues with previous results.
        \item Other interpretability measures.
        \begin{itemize}
            \item Open AI automated interpretability method and variations.
            \item Manual \url{https://transformer-circuits.pub/2023/monosemantic-features#appendix-rubric}.
            \item Feature sparsity.
        \end{itemize}
    \end{itemize}
\end{frame}

\end{document}